Convolutional Neural Network to Detect Equivalent Questions in Online Forums \\ DL-IC 2018 Project

### 2018-07-04 paper/paper_final.tex
paper update

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\subsubsection{The convolution}
We use a convolutional layer to compute the question-wide distributed vector representations $\vec{r}_{q1}$ and $\vec{r}_{q2}$. For each question, the convolutional layer first produces local features around each word in the question. Then, it combines these local features using a sum operation to create a fixed-sized feature vector (representation) for the question.\\
Given a question $q$, the convolutional layer applies a matrix-vector operation to each window of size $k$ of successive windows in $\vec{q}_{emb} = [\vec{r}^{w_1},\vec{r}^{w_2}, ..., \vec{r}^{w_N}]$. Let us define the vector $\vec{z}_n \in \mathbb{R}^{dk}$ as the concatenation of a sequence of $k$ word embeddings, centralized in the $n$-th word:
\begin{center}
$\vec{z}_n = [\vec{r}^{w_{n-(k-1)/2}},..., \vec{r}^{w_n}, ...,\vec{r}^{w_{n+(k-1)/2)}}]^T$
\end{center}•
The convolutional layer computes the $j$-th element of the vector $\vec{r}_{q1} \in \mathbb{R}^{cl_u}$ as follows: 
\begin{equation}
[\vec{r}_{q1}]_j = f \left( \sum_{1<n<N}[f(W^1 z_n + b^1)]_j \right)
\end{equation}•
where $W^1 \in \mathbb{R}^{cl_u \times dk}$ is the weight matrix of the convolutional layer and $f$ is the hyperbolic tangent function. The same matrix is used to extract local features around each word window of the given question. The global fixed-sized feature vector for the question is obtained by using the sum over all word windows.\\
Matrix $W^1$ and vector $b^1$ are parameters to be learned. The number of convolutional units $cl_u$ (which corresponds to the size of the question representation), and the size of the word context window $k$ are hyper-parameters to be tuned.

@@

%-------------------------------------------------------------------------
\subsection{Question of different length}
As starting point let's understand why unequal text length is a problem in NLP. The basic recipe for processing text in neural models is as follows:
\begin{itemize}\itemsep0.1pt
\item Convert text to tokens
\item Convert tokens to token IDs
\item Look up embedding vectors for IDs and construct a 2D tensor that are input to subsequent layers.
\end{itemize}•
Now, most neural network APIs (such as Keras) are “define-and-run” so we cannot change the dimensions of various matrices and tensors during runtime. All the techniques that can be used to tackle this issue force us to introduce new hyper-parameters or work-around in our model which are not fundamental for  the NLP task but are present to only handle unequal text lengths.\\
We decided to try two of these techniques, which we found out to be the best practices:
\begin{itemize}
	\item Fixing the length of all the questions to the \textbf{maximum} length. All the questions that are shorter are padded using the \emph{UNKNOWN} word (index $0$ of the vocabulary).
	\item Fixing the length of all the questions to the \textbf{mean} length. All the questions that are shorter are padded as in the case above; all the questions that are longer are chopped.
\end{itemize}
To keep the architecture of our network as simpler as possible we decided to apply this techniques as a preprocessing phase, without adding any further layer to the network architecture. Therefore, our network accepts as input only questions of the same length.\\
The experimental results show that the \emph{max} approach provides the best performances. This is probably due to the fact that the \emph{mean} approach is very likely to drop important semantic information from the sentence. It is worth mentioning that, in order for the \emph{max} approach to perform at its best, it can be a good choice dropping the outliers, i.e. the questions that are too long with respect to the other observations.
%-------------------------------------------------------------------------
\subsection{Type of convolution}

%-------------------------------------------------------------------------
\subsection{In-domain Word Embeddings}

>>>>

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{svg}

@@

\subsubsection{The convolution}
We use a convolutional layer to compute the question-wide distributed vector representations $\vec{r}_{q1}$ and $\vec{r}_{q2}$. For each question, the convolutional layer first produces local features around each word in the question. Then, it combines these local features using a sum operation to create a fixed-sized feature vector (representation) for the question.\\
Given a question $q$, the convolutional layer applies a matrix-vector operation to each window of size $k$ of successive windows in $\vec{q}_{emb} = [\vec{r}^{w_1},\vec{r}^{w_2}, ..., \vec{r}^{w_N}]$. Let us define the vector $\vec{z}_n \in \mathbb{R}^{dk}$ as the concatenation of a sequence of $k$ word embeddings, centralized in the $n$-th word:
\begin{center}
$\vec{z}_n = [\vec{r}^{w_{n-(k-1)/2}},..., \vec{r}^{w_n}, ...,\vec{r}^{w_{n+(k-1)/2)}}]^T$
\end{center}•
The convolutional layer computes the $j$-th element of the vector $\vec{r}_{q1} \in \mathbb{R}^{cl_u}$ as follows: 
\begin{equation}
[\vec{r}_{q1}]_j = f \left( \sum_{1<n<N}[f(W^1 z_n + b^1)]_j \right)
\label{eq:conv}
\end{equation}•
where $W^1 \in \mathbb{R}^{cl_u \times dk}$ is the weight matrix of the convolutional layer and $f$ is the hyperbolic tangent function. The same matrix is used to extract local features around each word window of the given question. The global fixed-sized feature vector for the question is obtained by using the sum over all word windows.\\
Matrix $W^1$ and vector $b^1$ are parameters to be learned. The number of convolutional units $cl_u$ (which corresponds to the size of the question representation), and the size of the word context window $k$ are hyper-parameters to be tuned.

@@

%-------------------------------------------------------------------------
\subsection{Question of different length}
As starting point let's understand why unequal text length is a problem in NLP. The basic recipe for processing text in neural models is as follows:
\begin{itemize}\itemsep0.1pt
\item Convert text to tokens
\item Convert tokens to token IDs
\item Look up embedding vectors for IDs and construct a 2D tensor that are input to subsequent layers.
\end{itemize}•
Now, most neural network APIs (such as Keras) are “define-and-run” so we cannot change the dimensions of various matrices and tensors during runtime. All the techniques that can be used to tackle this issue force us to introduce new hyper-parameters or work-around in our model which are not fundamental for  the NLP task but are present to only handle unequal text lengths.\\
We decided to try two of these techniques, which we found out to be the best practices:
\begin{itemize}
	\item Fixing the length of all the questions to the \textbf{maximum} length. All the questions that are shorter are padded using the \emph{UNKNOWN} word (index $0$ of the vocabulary).
	\item Fixing the length of all the questions to the \textbf{mean} length. All the questions that are shorter are padded as in the case above; all the questions that are longer are chopped.
\end{itemize}
To keep the architecture of our network as simpler as possible we decided to apply this techniques as a preprocessing phase, without adding any further layer to the network architecture. Therefore, our network accepts as input only questions of the same length.\\
The experimental results show that the \emph{max} approach provides the best performances. This is probably due to the fact that the \emph{mean} approach is very likely to drop important semantic information from the sentence. It is worth mentioning that, in order for the \emph{max} approach to perform at its best, it can be a good choice dropping the outliers, i.e. the questions that are too long with respect to the other observations.
%-------------------------------------------------------------------------
\subsection{Type of convolution}
To implement the convolution described in Section 3.1.2., we decided to use a 1-dimensional convolutional layer. The reason behind this choice is that, as images are seen as three input maps (R,G,B) where each map is a 2-dimensional array (matrix) with height and width, similarly text can be seen as $d$ input maps where each map has only one dimension which is the length of the question. This interpretation of the 1-dimensional convolution is exemplified in Figure \ref{fig:conv}.\\
After applying vertically the sum operation to the output of this layer, the resulting vector is exactly the one described by Equation \ref{eq:conv}

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.8\linewidth]{img/conv1d.jpg}
\end{center}
\caption{The 1-dimensional convolution.}
\label{fig:conv}
\end{figure}
%-------------------------------------------------------------------------
\subsection{In-domain Word Embeddings}
